Catchy Title Goes Here

### 2018-06-21 Paper/paper.tex
updated figures, added references, etc.

\section{Introduction \label{sec:Intro}}
Dramatic improvements in the coverage of vaccination with measles containing vaccines have lead to significant reductions in the incidence of measles disease and the associated childhood mortality [REF - Dabbagh et al]. 
However, despite these improvements at the national scale, significant heterogeneity in both vaccination coverage and disease incidence remains at the sub-national level [REF Takahashi].
Given the high transmission rate of measles, the proportion of the population that must be immunized to achieve and maintain elimination of measles is likely to be greater than 90\% in most populations [REF - Goodson?] (the exact level will be modulated by locally specific conditions just as popualtion density and contact rates).
%"The WHO recommends that programs use good quality data to monitor population immunity by identifying and responding to large numbers of susceptibles" from Winter et al reference to WHO Measles vaccines: WHO position paper Wkly Epidemiol Rec 2017;92:205-28
%"Because the risk of measles outbreaks is determined by the rate of accumulation of susceptible people in the population, programmes should use available good quality data on population immunity (i.e. vacci- nation coverage, surveillance, serological studies) to monitor the accumulation of susceptible people and conduct follow-up campaigns before the number of pre-school children susceptible to measles approaches the equivalent of one birth cohort, in order to prevent an outbreak of measles. This approach has been found to be programmatically useful in preventing large outbreaks. However, for large countries and countries which are close to measles elimination, a more exten- sive assessment of the accumulation of susceptible persons should be carried out at the subnational level." WHO Measles vaccines: WHO position paper Wkly Epidemiol Rec 2017;92:205-28 -- http://apps.who.int/iris/bitstream/handle/10665/255149/WER9217.pdf?sequence=1

The size and age-distribution of the population susceptible to measles (hereafter we refer to this as the "susceptible persons") depends on the competing prossess of susceptible recruitment through birth and susceptible loss through either natural infection or vaccination. 
Quantifying the size and age distribution of susceptible persons to measles is a critical tool in evaluating outbreak risk, the performance of vaccination programs, and developing vaccine-based interventions. 
The WHO recommends that programs monitor the accumulation susceptible persons using "good quality data" for all countries and that these should be made at the sub-national level for large countries and those close to measles elimination [REF -- Wkly Epid Rec].

Direct observation of the distribution of susceptible persons is challenging as it requires a sero-survey which can be cost-prohibitive (Winter et al JID paper). 
A sero-survey can provide a well resolved cross-sectional estimate of the distribution of suscpetible persons, it cannot, by itself, quantify the relative contribution of natural and vaccine derived immunity as current serological diagnostics cannot distinguish between these two sources [REF - Winter JID]; however, the relative contribution of these sources can be estimated through modeling [REF- Winter JID].
In the absence of a sero-survey, the number and age-distribution of susceptible persons can be estimated though demographic models that account for inputs through births and immunization via both vaccination an natural infection [Winter et al JID, Winter et al Madagascar, Merler et al http://www.thelancet.com/pdfs/journals/laninf/PIIS1473-3099(17)30421-8.pdf]. In this approach, the contribution of natural immunity is modeled as the result of a dynamic transmission model [REF] which requires explicit assumptions about the age-sepcific force of infection via a Who Acquires Infection From Whom matrix that describes the age-specific rate of infectious contacts from each age class to each other.  Direct estimation of this WAIFW matrix is challenging from disease incidence or seroprevalence data [though see Whittaker and Farrington etc.], and thus it is common to make simplfying assumptions that either the WAIFW matrix is constant for all ages (i.e. a uniformly well-mixed system) or that the WAIFW matrix is a scalar function of some other measureable interaction process, such as contacts measures via diary studies [REF -- Mossong POLYMOD study].

The catalytic model was initially proposed as a method for estimating the age-specific force of infection from cross-sectional age-specfic sero-prevalence data [REF - Griffiths]. 
This model, which represents the probability of immunization via natural infection at specific age $a$ as the cumulative sum of age-specific force of infection prior to age $a$, was then adapted by Grenfell and Anderson [REF - 1985] for use with cross-sectional, age-specific case observations.  
The resulting fitted model can then be used to estimate the age-specific distribution of susceptible persons.
Thus, in the absence of a sero-survey it is possible to estimate seroprevalence from routinely collected age-specific case reporting.

Original applications of the catalytic model assumed that natural infection was the only source of immunity [REF - Griffiths].  
In the vaccine era, however, the age distribution of susceptibility is generated by the sum of the rates of vaccination and natural infection. 
Though measles vaccination is recommended to follow a specific schedule, with children receiving two doses at prescribed ages, in practice, the ages at which receive an immunizing dose of measles containing vaccine (MCV) may be highly variable because of variaion in access to vaccination services [REF Metcalf; Takahashi; others] and the multiple vaccination initiatives that are employed: e.g. routine vaccination, supplementary immunization activities, child and maternal health days, school-based immunization drives, etc.
Recent work has highlighted significant variability both in the maximum vaccination coverage achieved and the timeliness of vaccination, that is, the proportion of children receiving vaccination at any given age.
Thus, vaccination can itself be modeled as an age-specific hazard rate that accounts for the sum of multiple forces.
Failing to account for the age-specific pattern of vaccination may lead to a biased interpretation of case-data, as children who might be vaccinated later than the recommended age remain susceptible may contribute to incident cases.

Further, vaccination does not necessarily imply immunization.  
The efficacy of measles vaccine is often assumed to be ... (see Uzicanin article) though is also known to improve with age as older children are more likely to have lost maternally transferred antibodies (se many). 
The effectiveness of vaccine delivered in field settings may also vary dramatically due to stability and effectiveness of the vaccine cold chain (Doshi et al 2017 Vaccine).  
A comparison of the age-distribution of vaccination and sero-prevalence may highlight areas with low effectiveness, and absent a sero-survey this assessment could be made using age-specific case records.

Here we present an extension of the catalytic model that represents the age-specific proportion immune as the cumulative sum of the hazards of both natural infection (force of infection) and effective immunization via vaccination.
We illustrate how this model can be used to estimate the age-specific sero-prevalence, and also the age-specific rates of vaccination and force of infection, and vaccine effectiveness using age-specific vaccine coverage data and case-records. 
We illustrate performance of this model using both simulated data, and measles case surveillance data from DRC combined with vaccination coverage surveys conducted as part of the 2013-14 DHS.  
A contemporary measles sero-survey conducted during the 2013-14 DHS allows us to validate the performance of our estimates of sero-prevelence against direct measurements. 
We finally present a fit of the model to the surveillance data, vaccine coverage data, and sero-survey and discuss opportunities for combining data sources.

@@

\subsection{Model Comparison\label{subsec:BF}}
For the real data from the DRC, we wish to determine which provinces warrant the more complicated seroprevalence model from Sec.~\ref{subsec:CompetingRates} that includes a constant in addition to the Weibull function, and which are described sufficiently by a constant force of infection alone. To do this, we calculate the Bayes factor between the two models, which, given that our prior belief in the two models is equal, is simply the ratio of the evidence for each of the models. The evidence is the fully marginalized likelihood (FML), and can often be quite difficult to calculate accurately. Luckily for us, the two models in question are nested - we recover the constant-only model when $\alpha = 1.0$ -  and so we can use a technique called the Savage-Dickey density ratio.

@@

\subsection{Simulated Data\label{subsec:SimData}}
To confirm that we can accurately recover the force of infection and vaccination hazard using the model we have chosen, we generate simulated datasets consisting of age-specified measles case, vaccination, and serology data using a previously developed, age-structured MSIRV (Maternally immune, Susceptible, Infected, Recovered, Vaccinated) model~\cite{Metcalf2012}. We used demographic parameters from UN estimates for the DRC and increased vaccination linearly from 0 to 50\% over the first 30 years of a 50 year simulation. $R_{0}$ (the number of secondary cases resulting from the introduction of a single infected individual) was assumed to be constant at 15 over the entirety of the simulation. The force of infection (foi) was not directly chosen to be a Wiebull function, but is generated using the specified $R_0$ and WAIFW (Who Acquires Infection from Whom) matrix that describes social interactions as estimated via the POLYMOD study~\cite{Mossong2008}. This means that we cannot directly compare the recovered Weibull parameters to injected parameters, but we \emph{can} compare the recovered foi curve to the true values, as well as the recovered vaccination efficacy ($\gamma$) to the true value.

@@

We also assess an additional measure of the `infection risk' inferred for each province, which we define as
\begin{equation}
r = \int_0^{10} \mathrm{foi}(a) da,
\end{equation}
 or the force of infection integrated from 0 to $10$ years.  Thus, this is an aggregate measure of the total infection risk up to age 10. We calculate this infection risk using  case and vaccination data only, and all three datasets. In addition we calculate this measure using the model fit to the serology data alone (without case or vaccination coverage data).  This latter measure reflects the estimate one would make if inference were based only on the serosurvey of children between the ages of 5-59 months.

@@

Using simulated age-specific case data and a simulated vaccine coverage survey, we estimated the cumulative proportion sero-positive as a function of age (Figure~\ref{fig:simplots}) under different scenarios of vaccine effectiveness (see definition above).  Using the model fit without using serology data, the estimated age-specific sero-prevalence was broadly consistent with the simualated sero-prevalence, though with a slight negative bias (Figure~\ref{fig:simplots} a). Similarly, the estimated cumulative probability of vaccination as a function of age was unbiased for the simulated values (Figure~\ref{fig:simplots} b).

The Weibull distribution used to approximate the age-specific FOI is an over-simplification of the complex age-specific mixing represented in the POLYMOD WAIFW matrix used in the MSIRV simulations; and is likely to be an over-simplification of the true, but unknown, age-specific FOI that exists in any real-world epidemiological setting.  Acknowledging that limitation, the estimated Weibull FOI broadly reflects the average age-specific FOI from the MSIRV model(Figure~\ref{fig:simplots} d) ; the latter is calculated as the product of the WAIFW matrix and the average age-specific vector of infectious cases over the full simulation.  The resulting Weibull fit, which is necessarily unimodal, tends to average over the two modes in the simulated FOI which reflect the effect of age-assortative mixing and interactions among generations (i.e. parents and children).  The estimated vaccination hazard clearly identifies the mode of age-specific vaccination and the magnitude of the vaccination hazard is positively correlated with the simulated values (Figure~\ref{fig:simplots} e).

The vaccine effectiveness, $\gamma$, which measures the likelihood that a child recorded as vaccinated in a vaccine survey was immunized by vaccine, is poorly estimated using a model fit only to age-specific case incidence and vaccination survey data (Figure~\ref{fig:simplots} f).  The fitted values of vaccine is positively correlated with the true value used in simulation, but exhibits significant negative bias; in simulation runs with low effectiveness, the resulting posterior estiamte was low and significantly different than the prior on $\gamma$ (Figure~\ref{fig:simplots} f).  This bias is off-set by slight biases in other parameters (see Supplement), which explains why the estimated age-specific sero-prevalence only exhibited a slight negative bias. Estimated vaccine effectiveness remains biased in the correspoding model that includes seroprevalence (Eq 8); though the correlation between the posterior mean estimate and the true value used in simulation is stronger and the variance of the posterior distribution is smaller when serology data are included (Figure~\ref{fig:simplots} g).  Thus, while vaccine effectiveness may always be challenging to estimate, the resulting estimate from this model may serve as an indicator of vaccine effectiveness.  Results for the age-specific sero-prevalence, probabilty of vaccination, and the hazards for FOI and vaccination (e.g. corresponding to (Figure~\ref{fig:simplots}) for the model fit to age-specific case incidence, vaccination survey data, and seroprevalence data are presented in the Supplement.

@@

For each of the 11 provinces in DRC, we fit both the constant and Weibull plus constant hazard models using age-specific case incidence and vacciation survey data, without and with the seroprevalence data from the 2014 DHS survey. When seroprevalence data were not included, the constant hazard model was selected as the best fit for Bas-Congo, Equateur, Kasai Occidental, Kasai Orientale, and Maniema.  When seroprevalence data were included in the model, the constant hazard model was selected as the best fit for Bas-Congo, Kasai Occidental, Kasai Orientale, and Maniema (FIG REF; note that panels in FIG REF 2 and FIG REF 3 are grouped into the provinces for which the Weibull plus constan, or constant hazard was selected as the best fit when all data were included in the model).  The age-specific seroprevalence estimated using only the case and vaccination coverage data tended to over-estimate the prevalence observed in the 2014 DHS and the estimate that resulted from the incluson of the serology data in the model (FIG REF). The magnitude of this bias varied from province to province; with the largest differences observed in Bandundu, Equateur, Nord Kivu, and Sud Kivu. In each of these cases, the model using only case and vaccination survey data would have given an over-estimate of the proportion immune at each age class, and thus an under-estimate of measles outbreak risk.

\begin{figure*}
\includegraphics[width=0.85\textwidth,angle=0]{figures/SeroPlots-crop.pdf}
\caption{\label{fig:sero} Left panel: color coding of locations used in our analysis. The data was split into the 11 provinces in existence between 1997 and 2015. The map shows these provinces in different colors, as well as the current administrative boundaries as black lines. The colors on this map correspond to the colors used for each province in all figures in this section. Right panel: seroprevalence data (points) and inferred values up to 25 years. The solid line shows the mean of the posterior distribution for the inferred seroprevalence using all three datasets, while the dashed lines show the full range of the posterior distribution generated using only vaccination and case data. The provinces before the blank space were fit using the Weibull plus constant model for force of infection, while those after (starting with Bas Congo) were fit using only a constant. For the provinces in which seroprevalence data makes a difference in the inferred values, the inferred curve is typically lower than that generated using only vaccination and case data. This is especially pronounced in Sud Kivu.}
\end{figure*}

@@

\begin{figure*}
\includegraphics[width=\textwidth,angle=0]{figures/VacFOIveff-crop.pdf}
\caption{\label{fig:foivaccveff} Left panel: Vaccination hazard (solid lines, right y-axis) and vaccination coverage (dashed lines, left y-axis) recovered using all three sets of data from all 11 provinces, generated from a set of posterior samples. For most provinces, the vaccination hazard is tightly peaked near one year, as expected. Middle panel: The recovered force of infection using all data (solid curve) and only vaccination and case data (dashed curves). As in Figure~\ref{fig:sero}, the solid curve is the mean of the posterior distribution, and the provinces before the blank space were fit using the Weibull plus constant model while those after were fit using the constant-only model for force of infection. In most cases, the inclusion of the serology data did not make a large difference in the inferred force of infection. Right panel: recovered value for vaccine efficacy ($\pm 1\sigma$) from analyses using all three data sets for each province. Because of our results from the simulated data, we do not expect that these values to reflect the true vaccine efficacy (in general, they are probably biased low), but we \emph{do} expect that the ranking of provinces by vaccine efficacy is correct.}
\end{figure*}

The resulting age-specific hazards for vaccination and FOI illustrate significant province-level variation.  Vaccination at the province-level exhibits strong differences in both the maximum coverage achieved and the timeliness (e.g. the variance of the vaccine hazard) (FIG REF a).  The hazard of vaccination falls to near 0 by approximately 24 months of age in all provinces except for Bas Congo, where the proportion of children vaccinated continues to increase over the full age range (6-59 months) included in the DHS survey (FIG REF a). In the 7 provinces for which the Weibull plus constant FOI hazard was selected, the mode of the fitted FOI was at or below 5 years of age (FIG REF b).  In 6 of 7 of these provinces, the FOI hazard rate dropped to near 0 by 10 years of age, suggesting little additional infection risk beyond 10 years of age.  In all 7 of these provinces, the fitted hazard rate for the model with and without serology data included were similar.  Equateur province stands out as an exception, here the FOI was highest of all provinces and remained high through adolescence. Notably, the mean age of infection among reported cases was highest in Equateur province. In the 4 provinces for which the contant hazard rate model was the best fit the rates were similar for models fit with and without the inclusion of serology data (FIG REF b).

There was high variability in the estimated vaccine effectiveness in the 11 provinces.  In 8 of 11 provinces, the fitted values are considerably lower than the conventionally assumed vaccine effciacy of ~85\% for measles vaccine [REF Uzacanin paper]. Based on the results from the simulation study, above, the absolute value of these estimates should be taken with caution; however, they may reflect a qualitative ranking of effectiveness, or the reliabilty of vaccination survey responses in these provinces.  The 4 provinces with the lowest values, Bandundu, Sud Kivu, Kasia Occidental, and Kasai Oriental, all reported relatively high vaccination coverage in the 2014 DHS relative to the seroprevalence.  Thus, low effectiveness is consistent with this observation.  Notably, this is corroborated by low estimated effectiveness when the model is fit in the absence of serological data.

@@

The age-specific profile of immunity to measles infection can be a powerful tool for the assessment of vaccination programs and the planning for vaccine-based interventions [REF - Winter; Cutts and Hanson]. Comprehensive serological surveys can be prohibitively expensive and are thus only conducted infrequently.  Age-specific measles case incidence, by contrast, is routinely collected in most national health systems and has been previously used to infer age-specific infection risk via the catalytic model [REF - Griffiths; Grenfell and Anderson; Whittaker and Farrington; Ferrari 2010]. Here we present a novel extension of the catalytic model that uses age-specific vaccination coverage surveys to account for the competing age-specific rates of immunization via vaccination and natural infection.  We show, using a simulation experiment, that this model can be used to estimate age-specific seroprevalence using age-specific case incidence and vaccination coverage data. A necessary component of this model is the vaccination effectiveness, which gives the proportion of individuals who report vaccination on a survey that are effectively immunized by vaccination.  We show that, while this vaccination effectiveness is difficult to estimate in absolute value, our model-based estimator is positively correalted with the true value used in simulation; thus this estimate may be useful as a relative measure of vaccination effectiveness.  We further show that the inclusion of limited serological survey data, here a survey of children between 6-59 months of age, results in improvements of all model parameters, including vaccination effectiveness.

>>>>

\section{Introduction \label{sec:Intro}}
Dramatic improvements in the coverage of vaccination with measles-containing vaccines have lead to significant reductions in the incidence of measles disease and the associated childhood mortality ~\cite{perry_global_2014}. Despite these improvements at the national scale, however, significant heterogeneity in both vaccination coverage and disease incidence remains at the sub-national level~\cite{prada_demographics_2017,takahashi_geography_2017,metcalf_transport_2015}.
Given the high transmission rate of measles, the proportion of the population that must be immunized to achieve and maintain elimination of measles is likely to be greater than 90\% in most populations [REF - Goodson?] (the exact level will be modulated by locally specific conditions such as population density and contact rates).
%"The WHO recommends that programs use good quality data to monitor population immunity by identifying and responding to large numbers of susceptibles" from Winter et al reference to WHO Measles vaccines: WHO position paper Wkly Epidemiol Rec 2017;92:205-28
%"Because the risk of measles outbreaks is determined by the rate of accumulation of susceptible people in the population, programmes should use available good quality data on population immunity (i.e. vacci- nation coverage, surveillance, serological studies) to monitor the accumulation of susceptible people and conduct follow-up campaigns before the number of pre-school children susceptible to measles approaches the equivalent of one birth cohort, in order to prevent an outbreak of measles. This approach has been found to be programmatically useful in preventing large outbreaks. However, for large countries and countries which are close to measles elimination, a more exten- sive assessment of the accumulation of susceptible persons should be carried out at the subnational level." WHO Measles vaccines: WHO position paper Wkly Epidemiol Rec 2017;92:205-28 -- http://apps.who.int/iris/bitstream/handle/10665/255149/WER9217.pdf?sequence=1

The size and age-distribution of the population susceptible to measles (hereafter we refer to this as the "susceptible persons") depends on the competing processes of susceptible recruitment through birth and susceptible loss through either natural infection or vaccination. 
Quantifying the size and age distribution of susceptible persons to measles is a critical tool in evaluating outbreak risk, the performance of vaccination programs, and developing vaccine-based interventions. The WHO recommends that programs monitor the accumulation of susceptible persons using "good quality data" for all countries, and that these should be made at the sub-national level for large countries and those close to measles elimination [REF -- Wkly Epid Rec].

Direct observation of the distribution of susceptible persons is challenging as it requires a sero-survey which can be cost-prohibitive. 
Moreover, while a sero-survey can provide a well resolved cross-sectional estimate of the distribution of susceptible persons, it cannot, by itself, quantify the relative contribution of natural and vaccine-derived immunity as current serological diagnostics cannot distinguish between these two sources~\ref{Winter2018Sero}. Fortunately, the relative contribution of these sources can be estimated through modeling.

In the absence of a sero-survey, the number and age-distribution of susceptible persons can be estimated though demographic models that account for inputs through births and immunization via both vaccination and natural infection~\cite{METCALF2017S14,TRENTINI20171089,Winter2018Sero}. In this approach, the contribution of natural immunity is modeled as the result of a dynamic transmission model [REF] which requires explicit assumptions about the age-specific force of infection via a Who Acquires Infection From Whom (WAIFW) matrix that describes the age-specific rate of infectious contacts from each age class to each of the others.  Direct estimation of this WAIFW matrix is challenging from disease incidence or seroprevalence data (though see Whittaker and Farrington etc.), and thus it is common to make simplifying assumptions that either the WAIFW matrix is constant for all ages (i.e. a uniformly well-mixed system) ,or that the WAIFW matrix is a scalar function of some other measurable interaction process, such as contacts measured via diary studies~\ref{Polymod}.

The catalytic model is another approach that was initially proposed as a method for estimating the age-specific force of infection from cross-sectional, age-specific sero-prevalence data~\cite{griffiths_catalytic_1974}. 
This model, which represents the probability of immunization via natural infection at a specific age, $a$, as the cumulative sum of age-specific force of infection prior to this age, was then adapted by Grenfell and Anderson~\cite{Grenfell1985} for use with cross-sectional, age-specific case observations.  
The resulting fitted model can then be used to estimate the age-specific distribution of susceptible persons.
Thus, in the absence of a sero-survey it is possible to estimate seroprevalence from routinely collected age-specific case reporting.

Original applications of the catalytic model assumed that natural infection was the only source of immunity~\cite{griffiths_catalytic_1974}.  
In the vaccine era, however, the age distribution of susceptibility is generated by the sum of the rates of vaccination and natural infection. 
Though measles vaccination is recommended to follow a specific schedule, with children receiving two doses at prescribed ages, in practice, the ages at which children receive an immunizing dose of measles containing vaccine (MCV) may be highly variable because of variation in access to vaccination services~\cite{METCALF2017S14,TRENTINI20171089,Winter2018Sero} and the multiple vaccination initiatives that are employed: e.g. routine vaccination, supplementary immunization activities, child and maternal health days, school-based immunization drives, etc.
Recent work has highlighted significant variability both in the maximum vaccination coverage achieved and the timeliness of vaccination - that is, the proportion of children receiving vaccination at any given age. Thus, vaccination can itself be modeled as an age-specific hazard rate that accounts for the sum of multiple forces.
Failing to account for the age-specific pattern of vaccination may lead to a biased interpretation of case-data, as children who might be vaccinated later than the recommended age remain susceptible and may contribute to incident cases.

An additional complication is that vaccination does not necessarily imply immunization.  
In developing-world settings, the efficacy of measles vaccine is often assumed to be $\sim 85 \%$ , though it is known to improve with age as older children are more likely to have lost maternally transferred antibodies~\cite{Uzicanin2011}. The effectiveness of vaccine delivered in field settings may also vary dramatically due to stability and effectiveness of the vaccine cold chain~\cite{Doshi2017}.  
A comparison of the age-distribution of vaccination and sero-prevalence may highlight areas with low effectiveness, and absent a sero-survey this assessment could be made using age-specific case records.

Here we present an extension of the catalytic model that represents the age-specific proportion immune as the cumulative sum of the hazards of both natural infection (force of infection) and effective immunization via vaccination.
We illustrate how this model can be used to estimate the age-specific sero-prevalence,  the age-specific rates of vaccination and force of infection, and vaccine effectiveness using age-specific vaccine coverage data and case-records. 
We illustrate performance of this model using both simulated data and measles case surveillance data from DRC combined with vaccination coverage surveys conducted as part of the 2013-14 DHS.  
A contemporary measles sero-survey conducted during the 2013-14 DHS allows us to validate the performance of our estimates of sero-prevelence against direct measurements. 
We finally present a fit of the model to the surveillance data, vaccine coverage data, and sero-survey and discuss opportunities for combining data sources.

@@

\subsection{Model Comparison\label{subsec:BF}}
For the real data from the DRC, we wish to determine which provinces warrant the more complicated seroprevalence model from Sec.~\ref{subsec:CompetingRates} that includes a constant in addition to the Weibull function, and which are described sufficiently by a constant force of infection alone. To do this, we calculate the Bayes factor (essentially the betting odds) between the two models, which, given that our prior belief in the two models is equal, is simply the ratio of the evidence for each of the models~\cite{TheRev}. The evidence is the fully marginalized likelihood (FML), and can often be quite difficult to calculate accurately. Luckily for us, the two models in question are nested - we recover the constant-only model when $\alpha = 1.0$ -  and so we can use a technique called the Savage-Dickey density ratio.

@@

\subsection{Simulated Data\label{subsec:SimData}}
To confirm that we can accurately recover the force of infection and vaccination hazard using the model we have chosen, we generate simulated datasets consisting of age-specified measles case, vaccination, and serology data using a previously developed, age-structured MSIRV (Maternally immune, Susceptible, Infected, Recovered, Vaccinated) model~\cite{Metcalf2012}. We used demographic parameters from UN estimates for the DRC and increased vaccination linearly from 0 to 50\% over the first 30 years of a 50 year simulation. $R_{0}$ (the number of secondary cases resulting from the introduction of a single infected individual) was assumed to be constant at 15 over the entirety of the simulation. The force of infection (foi) was not directly chosen to be a Wiebull function, but is generated using the specified $R_0$ and WAIFW (Who Acquires Infection from Whom) matrix that describes social interactions as estimated via the POLYMOD study~\cite{Polymod}. This means that we cannot directly compare the recovered Weibull parameters to injected parameters, but we \emph{can} compare the recovered foi curve to the true values, as well as the recovered vaccination efficacy ($\gamma$) to the true value.

@@

We also assess an additional measure of the `infection risk' inferred for each province, which we define as
\begin{equation}
r = \int_0^{10} \mathrm{foi}(a) da,
\end{equation}
 or the force of infection integrated from 0 to $10$ years.  Thus, this is an aggregate measure of the total infection risk up to age 10. We calculate this infection risk using  case and vaccination data only, and all three datasets. In addition we calculate this measure using the model analyez with the serology data alone (without case or vaccination coverage data).  This latter measure reflects the estimate one would make if inference were based only on the sero-survey of children between the ages of 5-59 months.

@@

Using simulated age-specific case data and a simulated vaccine coverage survey, we estimated the cumulative proportion sero-positive as a function of age (Figure~\ref{fig:simplots}) under different scenarios of vaccine effectiveness (see definition in Sec.~\ref{subsecSimData}).  Using the model analyzed without using serology data, the estimated age-specific sero-prevalence was broadly consistent with the simulated sero-prevalence, though with a slight negative bias (Figure~\ref{fig:simplots} (a) - top). Similarly, the estimated cumulative probability of vaccination as a function of age was unbiased for the simulated values (Figure~\ref{fig:simplots} (a) - middle).

The Weibull distribution used to approximate the age-specific FOI is an over-simplification of the complicated age-specific mixing represented in the POLYMOD WAIFW matrix used in the MSIRV simulations; and is likely to be an over-simplification of the true, but unknown, age-specific FOI that exists in any real-world epidemiological setting.  Acknowledging that limitation, the estimated Weibull FOI broadly reflects the average age-specific FOI from the MSIRV model(Figure~\ref{fig:simplots} (b) - top) ; the latter is calculated as the product of the WAIFW matrix and the average age-specific vector of infectious cases over the full simulation.  The resulting Weibull fit, which is necessarily unimodal, tends to average over the two modes in the simulated FOI which reflect the effect of age-assortative mixing and interactions among generations (i.e. parents and children).  The estimated vaccination hazard clearly identifies the mode of age-specific vaccination and the magnitude of the vaccination hazard is positively correlated with the simulated values (Figure~\ref{fig:simplots} (b) - bottom).

The vaccine effectiveness, $\gamma$, which measures the likelihood that a child recorded as vaccinated in a vaccine survey was immunized by vaccine, is poorly estimated using a model analyzed with only age-specific case incidence and vaccination survey data (Figure~\ref{fig:simplots} (c) - top).  The inferred values of vaccine efficacy are positively correlated with the true value used in simulation, but are biased significantly low; in simulation runs with low effectiveness, the resulting posterior estimate was low and significantly different than the prior on $\gamma$ (Figure~\ref{fig:simplots} f).  This bias is off-set by slight biases in other parameters (see Supplement), which explains why the estimated age-specific sero-prevalence only exhibited a slight negative bias. Estimated vaccine effectiveness remains biased in the corresponding model that includes seroprevalence (Eq 8); though the correlation between the posterior mean estimate and the true value used in simulation is stronger and the variance of the posterior distribution is smaller when serology data are included (Figure~\ref{fig:simplots} (c) - bottom).  Thus, while vaccine effectiveness may always be challenging to estimate, the resulting estimate from this model may serve as an indicator of vaccine effectiveness, and in particular the \emph{relative} values of vaccine effectiveness recovered for two different locations likely correctly identify the location with lower $\gamma$.  Results for the age-specific sero-prevalence, probability of vaccination, and the hazards for FOI and vaccination (e.g. corresponding to (Figure~\ref{fig:simplots}) for the model analyzed using age-specific case incidence, vaccination survey data, and seroprevalence data are presented in the Supplement.

@@

For each of the 11 provinces in DRC, we analyzed both the constant and Weibull plus constant hazard models using age-specific case incidence and vaccination survey data, first without and then with the seroprevalence data from the 2014 DHS survey. When seroprevalence data were not included, the constant hazard model was selected as the best model (via calculation of Bayes factors) for Bas-Congo, Equateur, Kasai Occidental, Kasai Orientale, and Maniema.  When seroprevalence data were included in the model, the constant hazard model was preferred for Bas-Congo, Kasai Occidental, Kasai Orientale, and Maniema.  The age-specific seroprevalence estimated using only the case and vaccination coverage data tended to over-estimate the prevalence observed in the 2014 DHS and the estimate that resulted from the inclusion of the serology data in the model~\ref{fig:sero}.(Note that in Fig.~\ref{fig:sero} (b) and ~\ref{fig:foivaccveff} (a) and (b), provinces are split by a blank space between those for which the Weibull plus constant model was preferred, and those for which the constant-only model was better). The magnitude of this bias varied from province to province, with the largest differences observed in Bandundu, Equateur, Nord Kivu, and Sud Kivu. In each of these cases, the model using only case and vaccination survey data would have given an over-estimate of the proportion immune at each age class, and thus an under-estimate of measles outbreak risk.

\begin{figure*}
\includegraphics[width=0.85\textwidth,angle=0]{figures/SeroPlots-crop.pdf}
\caption{\label{fig:sero} Panel (a): color coding of locations used in our analysis. The data was split into the 11 provinces in existence between 1997 and 2015. The map shows these provinces in different colors, as well as the current administrative boundaries as black lines. The colors on this map correspond to the colors used for each province in all figures in this section. Panel (b): seroprevalence data (points) and inferred values up to 25 years. The solid line shows the mean of the posterior distribution for the inferred seroprevalence using all three datasets, while the dashed lines show the full range of the posterior distribution generated using only vaccination and case data. The provinces before the blank space were analyzed using the Weibull plus constant model for force of infection, while those after (starting with Bas Congo) used only a constant. For the provinces in which seroprevalence data makes a difference in the inferred values, the inferred curve is typically lower than that generated using only vaccination and case data. This is especially pronounced in Sud Kivu.}
\end{figure*}

@@

\begin{figure*}
\includegraphics[width=\textwidth,angle=0]{figures/VacFOIveff-crop.pdf}
\caption{\label{fig:foivaccveff} Panel (a): Vaccination hazard (solid lines, right y-axis) and vaccination coverage (dashed lines, left y-axis) recovered using all three sets of data from all 11 provinces, generated from a set of posterior samples. For most provinces, the vaccination hazard is tightly peaked near one year, as expected. Panel (b): The recovered force of infection using all data (solid curve) and only vaccination and case data (dashed curves). As in Figure~\ref{fig:sero}, the solid curve is the mean of the posterior distribution, and the provinces before the blank space were fit using the Weibull plus constant model while those after were fit using the constant-only model for force of infection. In most cases, the inclusion of the serology data did not make a large difference in the inferred force of infection. Panel (c): recovered value for vaccine efficacy ($\pm 1\sigma$) from analyses using all three data sets for each province (solid), and vaccine and case data only (dashed). Note that the values without serology data are almost all consistent with the prior on $\gamma$. Because of our results from the simulated data, we do not expect that these values to reflect the true vaccine efficacy (in general, they are probably biased low), but we \emph{do} expect that the ranking of provinces by vaccine efficacy is correct.}
\end{figure*}

The resulting age-specific hazards for vaccination and FOI illustrate significant province-level variation.  Vaccination at the province-level exhibits strong differences in both the maximum coverage achieved and the timeliness (e.g. the variance of the vaccine hazard - is it tightly peaked or very broad?).  The hazard of vaccination falls to near 0 by approximately 24 months of age in all provinces except for Bas-Congo, where the proportion of children vaccinated continues to increase over the full age range (6-59 months) included in the DHS survey (see Fig.~\ref{fig:foivaccveff}(a)). In the 7 provinces for which the Weibull plus constant FOI hazard was selected, the mode of the fitted FOI was at or below 5 years of age - see Fig.~\ref{fig:foivaccveff}(b).  In 6 of 7 of these provinces, the FOI hazard rate dropped to near 0 by 10 years of age, suggesting little additional infection risk beyond 10 years of age.  In all 7 of these provinces, the fitted hazard rate for the model with and without serology data included were similar.  Equateur province stands out as an exception, here the FOI was highest of all provinces and remained high through adolescence. Notably, the mean age of infection among reported cases was highest in Equateur province. In the 4 provinces for which the constant hazard rate model was preferred, the rates were similar for models analyzed with and without the inclusion of serology data (Fig.~\ref{fig:foivaccveff}(b)).

There was high variability in the estimated vaccine effectiveness in the 11 provinces.  In 8 of 11 provinces, the fitted values are considerably lower than the conventionally assumed vaccine efficacy of ~85\% for measles vaccine~\cite{Uzicanin2011}. Based on the results from the simulation study, above, the absolute value of these estimates should be taken with caution; however, they may reflect a qualitative ranking of effectiveness, or the reliability of vaccination survey responses in these provinces.  The 4 provinces with the lowest values, Bandundu, Sud Kivu, Kasai Occidental, and Kasai Orientale, all reported relatively high vaccination coverage in the 2014 DHS relative to the seroprevalence.  Thus, low effectiveness is consistent with this observation.  Notably, this is corroborated by low estimated effectiveness when the model is analyzed in the absence of serological data  (Fig.~\ref{fig:foivaccveff}(c)).

@@

The age-specific profile of immunity to measles infection can be a powerful tool for the assessment of vaccination programs and the planning for vaccine-based interventions~\cite{CuttsHanson, Winter2018Sero}. Comprehensive serological surveys can be prohibitively expensive and are thus only conducted infrequently.  Age-specific measles case incidence, by contrast, is routinely collected in most national health systems and has been previously used to infer age-specific infection risk via the catalytic model~\cite{griffiths_catalytic_1974, Grenfell1985, farrington_correlated_2013, Ferrari2010FOI}. Here we present a novel extension of the catalytic model that uses age-specific vaccination coverage surveys to account for the competing age-specific rates of immunization via vaccination and natural infection.  We show, using a simulation experiment, that this model can be used to estimate age-specific seroprevalence using age-specific case incidence and vaccination coverage data. A necessary component of this model is the vaccination effectiveness, which gives the proportion of individuals who report vaccination on a survey that are effectively immunized by vaccination.  We show that, while this vaccination effectiveness is difficult to estimate in absolute value, our model-based estimator is positively correlated with the true value used in simulation; thus this estimate may be useful as a relative measure of vaccination effectiveness.  We further show that the inclusion of limited serological survey data, here a survey of children between 6-59 months of age, results in improvements of all model parameters, including vaccination effectiveness.
